---
title: "Brazilian GDP 2020:Q3"
description: "Data from the 2020:Q3 Brazilian GDP"
categories:
  - R
  - ggplot2
  - scraping
  - webscraping
  - reproducible-report
  - reproducible-research
  - PIB
  - GDP
author:
  - name: "Paulo Ferreira Naibert"
    url: https://github.com/pfnaibert/
repository_url: https://github.com/pfnaibert/pfnaibert.github.io
draft: true
date: 2020-12-03
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, error = TRUE, warning = TRUE)
knitr::opts_chunk$set(fig.width=8, fig.height=4)
```

#### Last Updated `r Sys.Date()`

In this post, we will do a not so deep dive in the IBGE release of the 2020:Q3 GDP results.
As usual, my source code for the functions can be found [here](https://github.com/pfnaibert/pfnaibert.github.io/blob/master/R/funs-macro.R).

First, let's load up our functions and other libraries.

```{r, error=T, warning=T, message=T}
# source functions
source("../../R/funs-gdp.R")
source("../../R/funs-gdp-floats.R")

# load libraries
library(kableExtra)
library(reshape2)
library(ggplot2)
```

Now, we load the data and make some transformations.

```{r}
# LOAD DATA

# levels
gdp.nominal   <- readRDS("../../data/gdp-nominal.rds")
gdp.real.NSA  <- readRDS("../../data/gdp-real-NSA.rds")
gdp.real.SA   <- readRDS("../../data/gdp-real-SA.rds")
gdp.index.NSA <- readRDS("../../data/gdp-index-NSA.rds")
gdp.index.SA  <- readRDS("../../data/gdp-index-SA.rds")

# rets
gdp.ret4    <- readRDS("../../data/gdp-ret4.rds")
gdp.ret1    <- readRDS("../../data/gdp-ret1.rds")
gdp.retsum4 <- readRDS("../../data/gdp-retsum4.rds")
gdp.retyear <- readRDS("../../data/gdp-retyear.rds")

# only gdp
gdp.rets <- cbind(gdp.ret4[, c(1,2,3)], "ret4"=gdp.ret4$GDP, "ret1"=gdp.ret1$GDP, "retsum4"=gdp.retsum4$GDP, "retyear"=gdp.retyear$GDP)

gdp.index <- cbind(gdp.index.NSA[, c(1,2,3)], "NSA"=gdp.index.NSA$GDP, "SA"=gdp.index.SA$GDP)
gdp.index$diff <- gdp.index$NSA - gdp.index$SA

# shares
# sum gdp.nom over 4 quarters and coerce to data.frame
gdp.nominal.ac4  <- data.frame( apply( gdp.nominal[,-c(1,2,3)], 2, sum4 ) )

# divide by gdp to get shares
gdp.share <- cbind( gdp.nominal[,c(1,2,3)], 100*gdp.nominal.ac4/gdp.nominal.ac4$GDP )
vab.share <- cbind( gdp.nominal[,c(1,2,3)], 100*gdp.nominal.ac4/gdp.nominal.ac4$VAB )
```

# Bird's Eye View

```{r}
gdp.fig.rets(gdp.ret4, 2007, 2, subtitle="T/T-4")
gdp.fig.rets(gdp.ret1, 2007, 2, subtitle="T/T-1")
gdp.fig.rets(gdp.retsum4, 2007, 2, subtitle="AC4Q T/T-4")
gdp.fig.rets(gdp.retyear, 2007, 2, subtitle="AC4Q T/T-4")
```

```{r}
gdp.tab.1(gdp.rets, "GDP Growths of the Last 5 Quarters")
```

## T/T-1 with Seasonal Adjustment

```{r, fig.width=8, fig.height=6.5}
gdp.ggplot.subsectors(gdp.ret1, "T/T-1 with Seasonal Adjustment")
```

```{r, fig.width=8, fig.height=5}
gdp.ggplot.demand.growth(gdp.ret1, subtitle = "T/T-1 with Seasonal Adjustment", 4)
```

```{r, fig.width=8, fig.height=5}
gdp.fig.index(gdp.index, 1996)
```

#### Trend Series
#### Seasonal Component Series

```{r}
gdp.tab.2(gdp.ret1, "GDP T/T-1 with Seasonal Adjustment")
```

## T/T-4

```{r, fig.width=8, fig.height=6.5}
gdp.ggplot.subsectors(gdp.ret4, "T/T-4")
```

```{r, fig.width=8, fig.height=5}
gdp.ggplot.demand.growth(gdp.ret4, subtitle = "T/T-4", 4)
```

```{r}
gdp.tab.2(gdp.ret4, "GDP T/T-4")
```

## T/T-4 AC4Q

```{r, fig.width=8, fig.height=6.5}
gdp.ggplot.subsectors(gdp.retsum4, "T/T-4 AC4Q")
```

```{r}
gdp.fig.rets.1(gdp.retsum4, 1997, qtr=2, subtitle="T/T-4 AC4Q")
```

```{r, fig.width=8, fig.height=5}
gdp.ggplot.demand.growth(gdp.retsum4, subtitle = "T/T-4 AC4Q", 2)
```

```{r}
gdp.tab.2(gdp.retsum4, "GDP T/T-4 AC4Q")
```

## Current Values


```{r}
gdp.tab.level(gdp.nominal, "GDP Levels by Activity Sector and Demand Variable (Millions of Current BRL)")
```

```{r}
gdp.tab.vab.share(vab.share, "VAB Shares")
```

```{r}
gdp.tab.demand.share(gdp.share, "GDP Shares by Demand Variables")
```

## Appendix

```{r}
# gdp.tab.2.full(gdp.ret4, "GDP T/T-4 Growths of the Last 5 Quarters")
```

## NEXT RELEASES FROM IBGE

|Reference Date | Publishing Date |
|:------------- |:--------------- |
| 2020:Q3       | 2020-12-03      |


## Useful Links
- [SIDRA-CNT](https://sidra.ibge.gov.br/home/cnt/brasil)
- [SCNT](https://www.ibge.gov.br/estatisticas/economicas/contas-nacionais/9300-contas-nacionais-trimestrais.html?=&t=o-que-e)
- [SCNT-releases](https://biblioteca.ibge.gov.br/index.php/biblioteca-catalogo?view=detalhes&id=72121)
- [2020:Q3 IBGE release](https://biblioteca.ibge.gov.br/visualizacao/periodicos/2121/cnt_2020_3tri.pdf)
- [2020:Q2 IBGE release](https://biblioteca.ibge.gov.br/visualizacao/periodicos/2121/cnt_2020_2tri.pdf)

# THANK YOU FOR READING!
