---
title: "Brazilian Seasonally Adjusted GDP"
description: "In this post, I will replicate the IBGE Seasonally Adjusted GDP from the GDP Chained Quarterly Index with the seasonal package"
categories:
  - R
  - ggplot2
  - PIB
  - GDP
  - seasonal
author:
  - name: "Paulo Ferreira Naibert"
    url: https://github.com/pfnaibert/
repository_url: https://github.com/pfnaibert/pfnaibert.github.io
draft: false
date: 2020-09-09
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, error = TRUE, warning = TRUE)
knitr::opts_chunk$set(fig.width=8, fig.height=4)
```

```{r, include=F}
plot.trend <- function(data)
{
newdata <- melt(data[, -c(2,3)], id="date")
idx2    <- (data$qtr==2 | data$qtr==4)
tmp     <- data$date[idx2]

ggplot(newdata, aes( x = date, y = value, color=variable) ) +
theme_bw() +
geom_line(size=1 ) +
scale_color_manual(values=c("blue")) +
scale_x_continuous(labels=date2qtr(tmp), breaks=tmp) +
scale_y_continuous(limits = c(90, NA), breaks=seq(90, 180, by=10) ) +
theme(legend.position="none",
	  legend.title=element_blank(),
	  axis.text.x = element_text(color = "black", size = 9, angle=90, vjust=.5),
	  # panel.grid.major.x = element_blank(),
	  panel.grid.minor.x = element_blank(),
  	  plot.title = element_text(face = "bold", hjust = .5),
  	  plot.subtitle = element_text(hjust = .5),
  	  plot.caption = element_text(size=11, hjust = 0) ) +
xlab("") + ylab("") +
labs(title="GDP Chained Quarterly Index",
 	 subtitle="Trend Component Series",
	 caption="Data Source: IBGE.")
}
```

```{r, include=F}
plot.cycle <- function(data)
{
newdata <- melt(data[, -c(2,3)], id="date")
idx2    <- (data$qtr==2 | data$qtr==4)
tmp     <- data$date[idx2]

ggplot(newdata, aes( x = date, y = value, color=variable) ) +
theme_bw() +
geom_line(size=1 ) +
scale_color_manual(values=c("blue")) +
scale_x_continuous(labels=date2qtr(tmp), breaks=tmp) +
scale_y_continuous(limits = c(95, 105), breaks=seq(95, 105, by=5) ) +
theme(legend.position="none",
	  legend.title=element_blank(),
	  axis.text.x = element_text(color = "black", size = 9, angle=90, vjust=.5),
	  # panel.grid.major.x = element_blank(),
	  panel.grid.minor.x = element_blank(),
  	  plot.title = element_text(face = "bold", hjust = .5),
  	  plot.subtitle = element_text(hjust = .5),
  	  plot.caption = element_text(size=11, hjust = 0) ) +
xlab("") + ylab("") +
labs(title="GDP Chained Quarterly Index",
 	 subtitle="Seasonal Component Series",
	 caption="Data Source: IBGE.")
}
```

```{r, include=F}
myplot <- function(data, year=1996)
{
newdata <- melt(data[, -c(2,3)], id="date")
idx2    <- (data$year >= year & (data$qtr==2 | data$qtr==4) )
tmp     <- data$date[idx2]

ggplot(newdata, aes( x = date, y = value, color=variable) ) +
theme_bw() +
geom_line(size=1 ) +
geom_hline(yintercept=tail(data$NSA,1), linetype="dashed") +
scale_color_manual(values=c("blue", "red"), labels=c("PIB Observado", "PIB Ajustado")) +
scale_x_continuous(labels=date2qtr(tmp), breaks=tmp) +
scale_y_continuous(limits = c(70, NA), breaks=seq(70, 180, by=10) ) +
theme(legend.position = "bottom",
	  legend.title=element_blank(),
	  axis.text.x = element_text(color = "black", size = 9, angle=90, vjust=.5),
	  # panel.grid.major.x = element_blank(),
	  panel.grid.minor.x = element_blank(),
  	  plot.title = element_text(face = "bold", hjust = .5),
  	  plot.caption = element_text(size=11, hjust = 0) ) +
xlab("") + ylab("") +
labs(title="GDP Chained Quarterly Index, 1995=100",
	 caption="Data Source: IBGE.")
}
```

#### Last Updated `r Sys.Date()`

On the IBGE publication of the Brazilian GDP data, they release, among others, 3 series:

1. GDP Chained Quarterly Index: Table 1620
2. GDP Chained Quarterly Index with Seasonal Adjustment: Table 1621
3. GDP Growth T/T-1 with Seasonal Adjustment: Table 5932

On this post, I will try to match the IBGE model of Seasonally Adjusted using the X13-ARIMA software.
A note of caution is in order here, I am not actually trying to model the series, what we are trying to do is to match the IBGE model.

To replicate the IBGE model, we use the GDP Chained Quarterly Index (with NO seasonal adjustment) and use the X13-ARIMA software through the [`seasonal`](https://cran.r-project.org/web/packages/seasonal/index.html) package.

### Installation

The `seasonal` package depends on the [`x13binary`](https://cran.r-project.org/web/packages/x13binary/index.html) package.
It should automatically download and compile the `x13` program.
But, alas, it didn't work on my machine, so I manually downloaded the program from the [census bureau website](https://www.census.gov/srd/www/x13as/) and indicated to `R` where the program was located.

Let's load the `seasonal` package and use the `checkX13()` function to check if it is working:

```{r, error=T, warning=T, message=T}
library(seasonal)

checkX13()
```

OK. It works.

## Let's get to Work

First, let's load up our functions and other libraries.

```{r, error=T, warning=T, message=T}
# source functions
source("../../R/funs-gdp.R")

# load libraries
library(reshape2)
library(ggplot2)
```

Now, we load the data and make some transformations.

```{r}
# LOAD DATA
gdp.index.NSA <-  readRDS("../../data/gdp-index-NSA.rds")
gdp.index.SA  <-  readRDS("../../data/gdp-index-SA.rds")

gdp.nsa <- df2ts( gdp.index.NSA )[, "GDP"]
gdp.sa1 <- df2ts( gdp.index.SA )[, "GDP"]

gdp.index <- cbind(gdp.index.NSA[, c(1,2,3)], "NSA"=gdp.index.NSA$GDP, "SA"=gdp.index.SA$GDP)
```

Now that the data is loaded, let's make a plot.


```{r, fig.width=8, fig.height=5, echo = F}
myplot(gdp.index)
```

## X13 Arima

### IBGE Model
On the [IBGE release](https://biblioteca.ibge.gov.br/visualizacao/periodicos/2121/cnt_2020_2tri.pdf) we can find the model used by IBGE on the Table on Page 24.
There it says:

|Variável |  sazonalidade | decomposição   | modelo arima   | efeitos de intervenção            |
|:------- |:------------- |:-------------- |:---------------|:--------------------------------- |
|PIB      |  Sim          | multiplicativo | (2 1 2)(0 1 1) | TD, Easter[1], LS2008.4, AO2020.2 |

So let's try to replicate it using the `seasonal` package.
The X13-ARIMA should do everything automatically.

## Trying (and Failing) to replicate

### Automatic Model

First, we try a vanilla call, just `seas(gdp.nsa)`, let's see what it outputs:

```{r}
# Vanilla Model
m0 <- seas(gdp.nsa)
summary(m0)
```

Also, we can check the call used by the function:

```{r}
static(m0)
```

OK, so it is a SARIMA model $(0, 1, 0)(0, 1, 1)$ with level shift on 2008:Q4 and an outlier on 2020:Q2.
Also, there is a weekday effect.
Comparing with the IBGE model, it is NOT the same.
The IBGE model has SARIMA of order $(2, 1, 2)(0, 1, 1)$ and easter effect.
About the Trading Days (TD), the IBGE release is a little mysterious, what kind of TD did they use?
There is not a unique specification for that effect.

Now, let's check the difference between IBGE's seasonal ajustment and ours.

```{r}
mean( abs( gdp.sa1 - final(m0) ) )
all.equal(gdp.sa1, final(m0))
```

So it is *really not* the model we want.

### Manual SARIMA (2 1 2)(0 1 1) Model

Let's try some more specifications:

```{r}
m1 <- seas(
x = gdp.nsa,
transform.function = "log",
arima.model = "(2 1 2)(0 1 1)",
)
summary(m1)
mean( abs( gdp.sa1 - final(m1) ) )
all.equal(gdp.sa1, final(m1))
```

OK, there's something wrong.
Checking the [IBGE CNT website](https://www.ibge.gov.br/estatisticas/economicas/contas-nacionais/9300-contas-nacionais-trimestrais.html?=&t=downloads) under `Ajuste_Sazonal` there is a pdf that explains how the function call is made by the IBGE researcher.
We can use the `view(m1)` function to get direct access to the `X13` program by means of a `shiny` app.
On the `X13` program we have to use the options listed below (the IBGE pdf has a typo, it says *tipes* when it should be *types*):

```
transform{
  function = auto
}

regression{
  aictest = (td easter)
}

pickmdl{
  method = best
  identify = all
}

outlier{
  types = all
}

forecast{
  maxlead = 6
  maxback = 0
}

x11{
  savelog = q
}
```

Sending the output to the console, we get the `static()` call, which can be summarised as:


```{r}
m1 <- seas(x = gdp.nsa,
	 automdl = NULL,
	 pickmdl.method = "best",
	 pickmdl.identify = "all",
	 outlier.types = "all",
	 forecast.maxlead = 6,
	 forecast.maxback = 0,
	 x11.savelog = "q")
```


## The IBGE specification

```{r}
summary(m1)
mean(abs( gdp.sa1 - final(m1) ))
all.equal(gdp.sa1, final(m1))
static(m1)
```

Now, not only the function returned the specification described on the [IBGE release](https://biblioteca.ibge.gov.br/visualizacao/periodicos/2121/cnt_2020_2tri.pdf), but also the differences are after the 4th decimal place (which suggests rounding errors only).

### Seasonally Adjusted Series Plot

Let's plot the Seasonally Adjusted Index and the differences of the IBGE original series from Table 1621 and our X13 adjustment:

```{r}
plot(final(m1), col = "blue", lwd=2)
lines(gdp.sa1, col="red", lwd=2)

plot(final(m1)-gdp.sa1)
abline(h=0)
```

Yeah, it is pretty close.

## T/T-1 Growth

Also, let's check if the T/T-1 returns match:

```{r}
gdp.ret1.full <-  readRDS("../../data/gdp-ret1.rds")
gdp.ret <- df2ts( gdp.ret1.full )[, "GDP"]

myret1 <- ts(round(ret1(gdp.sa1),1), start=c(1996,1), freq=4)
myret2 <- ts(round(ret1(final(m1)),1), start=c(1996,1), freq=4)
all.equal(gdp.ret, myret1)
all.equal(gdp.ret, myret2)
```

It does! We did it!

## More Plots

Finally, let's replicate the Figures I.4 and I.5 of the IBGE release with ggplot:

```{r, fig.width=8, fig.height=5, echo = F}
tmp <- cbind(gdp.index.NSA[, c(1,2,3)], "trend"=trend(m1))
plot.trend(tmp)
```

```{r, fig.width=8, fig.height=5, echo = F}
tmp <- cbind(gdp.index.NSA[, c(1,2,3)], "cycle"=100*series(m1, "d10"))
plot.cycle(tmp)
```

And it looks a lot like they are the same plots.

## Final Thoughts

So, thanks to the IBGE documentation we could replicate the Seasonally Adjusted series.
I think I only wished it was a little easier to find it; also in the release, there could be calls made to the X13 program.
But well, at least they did publish the calls, it was just a bit out of the way for us to find it.

# THANK YOU FOR READING!
