---
title: "Graphing the Brazilian GDP in `R`"
description: Graphing the Brazilian GDP in `R`.
categories:
  - R
  - R-base
  - GDP
  - PIB
  - plot
author:
  - name: "Paulo Ferreira Naibert"
    url: https://github.com/pfnaibert
date: "2020-08-17"
preview: gdp.png
draft: false
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

#### Updated 2020-09-03

In this post, we want to make a similar plot as the figure in
[this pdf](https://portalibre.fgv.br/sites/default/files/2020-06/comunicado-do-comite-de-datacao-de-ciclos-economicos-29_06_2020-1.pdf).

# Import

First, I already downloaded the data and saved the only column that I will need in a `.csv` file.
Next, I will import the data to `R` and get some quick visualizations.

```{r}
# Import
PIB <- read.csv("../../data/PIB_BASE_2000.csv")

# info about PIB
str(PIB)

# quick visualizations
head(PIB)
tail(PIB)
```

We can see that we don't need the last two lines and that the first row are the dates.
(Yes, in `PT-BR` *date* is called *data*; also, in case you are wondering, *data* is called *dados*, which is just the latin translation. Anyway, moving on).
We are going to save the dates in vector in `dates` and delete the first column.

```{r}
# trim the last two lines
PIB <- head(PIB, -2)

# get first row and name it "dates"
dates <- PIB[,1]

# delete the first column
PIB <- PIB[,-1]
```

After deleting the first column, we can see that `PIB` changed class:

```{r}
# PIB is a different class now
str(PIB)
```

Before we give names to the values, let's format it more to our liking.

```{r}
# rearrange strings in dates
head(dates)
dates <- gsub(" T", ":Q", dates)
head(dates)
```

Note that `PIB` is a `numeric`, so we put `names`, not `rownames`.

```{r}
# give names to values and rearrange string
names(PIB) <- dates

# visualize again
head(PIB)
tail(PIB)
```

Now we can plot our data.

```{r}
plot(PIB, t="l")
```

This plot doesn't have the correct x labels.
We will correct that by using a `ts` object.

# `ts` object

It is pretty straightforward to make a `ts` object.

```{r}
PIB.ts <- ts(PIB, start=c(1980, 1), end=c(2014, 3), freq=4 )
```

Now, we plot `PIB.ts`.

```{r}
# graph it
plot(PIB.ts)
```
And the x labels are correct.

# Recession Dates

Now we get the dates to put the recession bars in the graph (yes, **manually**).

```{r}
# get recession limits in dates
recs <- c("1981:Q1", "1983:Q1", "1987:Q3", "1988:Q3", "1989:Q3", "1992:Q1", "1995:Q2", "1995:Q3", "1998:Q1", "1999:Q1", "2001:Q2", "2001:Q4", "2003:Q1", "2003:Q2", "2008:Q4", "2009:Q1" )
```

And we can make a nicer display with `kable()`.

```{r}
# TABLE
tab.rec <- t( matrix( recs, nrow=2 ) )
colnames(tab.rec) <- c( "begin", "end")
kable(tab.rec)
```

The dates in `recs` are correct and are more intuitive, however, `rect()` and `abline()` cannot understand that format, so we will use numeric dates.

```{r}
# get recession limits in numeric dates
qtr2num <- function(data) return( as.numeric(substring(data, 1, 4)) - 1/4 + as.numeric( substring(data, 7) )/4 )
recs2 <- qtr2num(recs)
recs2
```

Or we can get the series indices of the recession dates.

```{r}
# get recession limits in indices
idx <- unlist( sapply(recs, function(x) which(x==dates) ) )
names(idx) <- recs2; idx
```

Now that we have the recession limit dates, we plot the data with some more options, and use the `rect()` function to plot the graph with the recession bars.

```{r}
# graph it
plot(PIB.ts, xlim = c(1980, 2015), ylim=c(60,200), col="blue", lwd=1.5, main ="Cronologia Trimestral dos Ciclos de Negócios Brasileiros", xlab = "", ylab="")
for(i in 1:NROW(tab.rec)) rect(recs2[1+2*(i-1)], 1, recs2[2*i], 250, col="gray", border=T)
```

Wellp, that didn't go according to plan!


OK, we will use `plot(NULL, ...)` to remedy the situation:

```{r}
# graph it
plot(NULL, xlim = c(1980, 2015), ylim=c(60,200), main ="Cronologia Trimestral dos Ciclos de Negócios Brasileiros", xlab = "", ylab="")
for(i in 1:NROW(tab.rec)) rect(recs2[1+2*(i-1)], 1, recs2[2*i], 250, col="gray", border=T)
lines(PIB.ts, ylim=c(60,200), col="blue", lwd=1.5)
```

# Final Remarks
We did it!

Not quite.
Our data is not Seasonally Adjusted, and our series ends in 2014:Q3, while the orginal graphs ends in 2020:Q1.
However, this is just a matter of getting the data and plugging in the functions.
For **now**, our work is done.

In the next post I intend to make wrappers of the functions used to automate the whole thing.

