---
title: "Interactive Plots with `dygraphs`"
description: "`dygraphs` allows us to make interactive plots in webpages."
categories:
  - R
  - dygraphs
  - PIB
  - GDP
  - plot
author:
  - name: "Paulo Ferreira Naibert"
    url: https://github.com/pfnaibert/
repository_url: https://github.com/pfnaibert/pfnaibert.github.io
date: "2020-08-21"
draft: false
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
knitr::opts_chunk$set(fig.width=8, fig.height=4)
```

On previous posts, I showed how to make plots with the `base` and `ggplot2` libraries.
Today, I will make **interactive** plots with dygraphs.
This package has a terrific documentation website [here](https://rstudio.github.io/dygraphs/).

First, let's load the data and the functions:

```{r}
# import funs
source("../../R/funs.R")

# import data
y     <- import_gdp("../../data/PIB_BASE_2000.csv")

# visualization
head(y)
```


Now, we load `dygraphs` and try to make a plot:

```{r}
# libraries
library(dygraphs)

# plot
dygraph(y, main = "Cronologia Trimestral dos Ciclos de Negócios Brasileiros")
```

OH! That worked out of the box, with similar syntax as `base::plot()`.
Take that `ggplot`!

Ok, let's try some options.
First, let's label the series as "GDP".

```{r}
dygraph(y, main = "Cronologia Trimestral dos Ciclos de Negócios Brasileiros") %>%
dySeries(label = "GDP") %>%
dyAxis("x", drawGrid = FALSE)
```

Yes, we have to use the pipe `%>%`, so not EVERYTHING is similar to the `base::plot()`.
As I said about `ggplot`, this is a fancy package.
Let's have go at the range selector option.

```{r}
dygraph(y, main = "Cronologia Trimestral dos Ciclos de Negócios Brasileiros") %>%
dySeries(label = "GDP") %>%
dyAxis("x", drawGrid = FALSE) %>%
dyRangeSelector(height = 20)
```

COOL!
Now, let's try to make lines.
`dygraphs` uses `dyEvent()` for vertical lines and `dyLimit()` for horizontal lines.

```{r}
dygraph(y, main = "Cronologia Trimestral dos Ciclos de Negócios Brasileiros") %>%
dySeries(label = "GDP") %>%
dyAxis("x", drawGrid = FALSE) %>%
dyEvent("1983-03-30", "LABEL",  labelLoc = "top") %>%
dyEvent("2010-03-30", "LABEL",  labelLoc = "bottom") %>%
dyLimit( as.numeric(100), "LABEL", labelLoc = "right", color="red" )
```

So far, so good, now, let's make those rectangles.
`dygraphs` uses `dyShading()` to make those.
Before, we have to load the recession limit dates.

```{r}
# load recs
recs  <- qtr2date( readRDS("../../data/recessions.rds") )

# visualization
head(recs)

# tranform recs to data frame
recs.mat <- t( matrix( recs, nrow=2 ) ); colnames(recs.mat) <- c("start", "end"); recs.mat

# plot
dygraph(y, main = "Cronologia Trimestral dos Ciclos de Negócios Brasileiros") %>%
dySeries(label = "GDP") %>%
dyAxis("x", drawGrid = FALSE) %>%
dyShading( recs.mat[1, 1], recs.mat[1, 2] ) %>%
dyShading( recs.mat[2, 1], recs.mat[2, 2] ) %>%
dyShading( recs.mat[3, 1], recs.mat[3, 2] ) %>%
dyShading( recs.mat[4, 1], recs.mat[4, 2] ) %>%
dyShading( recs.mat[5, 1], recs.mat[5, 2] ) %>%
dyShading( recs.mat[6, 1], recs.mat[6, 2] ) %>%
dyShading( recs.mat[7, 1], recs.mat[7, 2] ) %>%
dyShading( recs.mat[8, 1], recs.mat[8, 2] )
```

NICE!
But it seems a lot of `dyShading` lines, can we make it like in `ggplot`?

```{r, error=T}
dygraph(y, main = "Cronologia Trimestral dos Ciclos de Negócios Brasileiros") %>%
dySeries(label = "GDP") %>%
dyAxis("x", drawGrid = FALSE) %>%
dyShading( recs.mat)
```

OH! Do I have to provide two arguments?

```{r, error=T}
dygraph(y, main = "Cronologia Trimestral dos Ciclos de Negócios Brasileiros") %>%
dySeries(label = "GDP") %>%
dyAxis("x", drawGrid = FALSE) %>%
dyShading( recs.mat, recs.mat )
```

NOPE!
Should I define the dimensions?

```{r, error=T}
dygraph(y, main = "Cronologia Trimestral dos Ciclos de Negócios Brasileiros") %>%
dySeries(label = "GDP") %>%
dyAxis("x", drawGrid = FALSE) %>%
dyShading( recs.mat[, 1], recs.mat[, 2] )
```

No, again.
Oh well, should I try a `for` loop?

```{r, error=T}
nrecs <- length(recs)/2; nrecs

dygraph(y, main = "Cronologia Trimestral dos Ciclos de Negócios Brasileiros") %>%
dySeries(label = "GDP") %>%
dyAxis("x", drawGrid = FALSE) %>%
for(i in 1:nrecs) dyShading( recs.mat[i, 1], recs.mat[i, 2] ) %>%
```

$%&\*# !!!

OK, calm down, maybe it's the pipe in the end, let's change the order of the lines:

```{r, error=T}
dygraph(y, main = "Cronologia Trimestral dos Ciclos de Negócios Brasileiros") %>%
dySeries(label = "GDP") %>%
for(i in 1:nrecs) dyShading( recs.mat[i, 1], recs.mat[i, 2] ) %>%
dyAxis("x", drawGrid = FALSE)
```

Well, it wasn't meant to be.
But hey, the `dyShading` feature works, I just don't know how to automate them.
If you had more patience than me, give me some feedback on the comments.

Thank you for reading.
