---
title: "Times Series Plots with `ggplot2`"
description: "`ggplot2` enables us to make beautiful graphics (as long as you use the tidyverse framework)"
categories:
  - R
  - ggplot2
  - PIB
  - GDP
  - plot
author: "Paulo Ferreira Naibert"
date: "2020-08-19"
preview: economist_plot.png
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
knitr::opts_chunk$set(fig.width=sqrt(2)*5, fig.height=5)
```

On previous posts
([here](https://pfnaibert.github.io/posts/2020-08-17-plot-base-gdp/) and [here](https://pfnaibert.github.io/posts/2020-08-18-wrapper-gdp/) ),
I demonstrated how to import and graph data using `R`.
On this post, I will make those same plots using `ggplot2` to make *beautiful* plots.
First we will need to load the data and the functions.

```{r}
# funs
source("../../_resources/funs.R")

# import data
y    <- import_gdp("../../_resources/data/PIB_BASE_2000.csv")
recs <- readRDS("../../_resources/data/recessions.rds")

# transform dates
dates <- date2num(names(y))
recs  <- date2num(recs)

# quick visualizations
head(y); head(dates); head(recs)
```

Now, let's load the `ggplot2` library and make some graphs.

```{r, error = T}
# libraries
library(ggplot2)

# expected plot
ggplot(y)
```

Wait, WHAT? Why didn't that work? I mean, `plot(y)`, should work:

```{r}
# base plot
plot(y)
```

And it DOES, with the labels working properly and we don't even have to specify the type (which should be line).
So why `ggplot2` does not?
Let's read the error message, as we should have, from the beginning.
OK, OK, We will have to convert our data to a `data.frame`, a bit cumbersome, but hey, this is a FANCY package.

```{r}
y.df  <- as.data.frame(cbind(dates, "GDP"=y) )
head(y.df)

# plot
ggplot(y.df)
```

It is just a BLANK plot?! OMG! Let's just
[RTFM](https://en.wikipedia.org/wiki/RTFM)
already and get it over with.

I used the following links (because internet is better than books?):

- http://www.cookbook-r.com/Graphs/
- https://www.tutorialspoint.com/ggplot2/ggplot2_time_series.htm
- https://www.neonscience.org/dc-time-series-plot-ggplot-r

So, `ggplot2` has its very own syntax [(grammar?)](https://ggplot2.tidyverse.org/), but it does have a quick plot option, let's try it now:

```{r}
# quick plot
qplot(x=dates, y=GDP, data=y.df,
	  geom="line", size=I(1), color = I("blue"),
	  xlab="", ylab="",
	  main = "Cronologia Trimestral dos Ciclos de Negócios Brasileiros" )
```

OK, that's a nice enought looking plot, let's try the `ggplot` function (again, the package has it's own grammar):

```{r}
# plot
ggplot(data = y.df, aes( x = dates, y = GDP ) ) +
	theme( plot.title = element_text(hjust = 0.5, face = "bold" ) ) +
	ggtitle( "Cronologia Trimestral dos Ciclos de Negócios Brasileiros" ) +
	xlab("") + ylab("") +
   	geom_line( color = "blue", size=1 )
```

For the title options I used the following links:

- https://stackoverflow.com/questions/40675778/center-plot-title-in-ggplot2
- http://www.cookbook-r.com/Graphs/Titles_(ggplot2)/

## Figures inside the plot

Cool, but what about `abline()` and `rect()`?

### `abline()`
For straigh lines in the plot, the `ggplot()` syntax is actually pretty intuitive, below we show an example:

```{r}
ggplot(data = y.df, aes( x = dates, y = GDP ) ) +
	theme( plot.title = element_text(hjust = 0.5, face = "bold" ) ) +
	ggtitle( "Cronologia Trimestral dos Ciclos de Negócios Brasileiros" ) +
	xlab("") + ylab("") +
	geom_line( color = "blue", size=1 ) +
	geom_vline( color = "red", linetype="dotted", size=.5, aes( xintercept=1990 ) ) +
	geom_hline( color = "red", linetype="dashed", size=.5, aes( yintercept=100 ) )
```

### `rect()`

NICE! Now I want to make a rectangle to highlight the recession periods.
The plot worked just fine when I used a single rectangle, but it would not work if I used a `for` loop to make more rectangles.
To figure out how to make more rectangles, I used [this page]( https://plotly.com/ggplot2/geom_rect/).
Let's show our work.

```{r}
# tranform recs to data frame
recs.mat <- as.data.frame( t( matrix( recs, nrow=2 ) ) ); colnames(recs.mat) <- c("start", "end")

# plot with rect
ggplot(data = y.df, aes( x = dates, y = GDP ) ) +
	theme( plot.title = element_text(hjust = 0.5, face = "bold" ) ) +
	ggtitle( "Cronologia Trimestral dos Ciclos de Negócios Brasileiros" ) +
	xlab("") + ylab("") +
	geom_rect( data=recs.mat, aes( NULL, NULL, xmin = start, xmax = end ), ymin = 0, ymax = 200 ) +
   	geom_line( color = "blue", size=1 )
```

And there you have it, I made the plot with `ggplot()`.

## Plot Objects

One neat thing about `ggplot()` is that we can "save" a plot as an object.
We can use this to customize the plot later.

```{r}
# save plot
gdp_plot  <-
ggplot(data = y.df, aes( x = dates, y = GDP ) ) +
	ggtitle( "Cronologia Trimestral dos Ciclos de Negócios Brasileiros" ) +
	xlab("") + ylab("") +
	geom_line( color = "blue", size=1 )

# list objects
ls()
```

Let's plot it.

```{r}
# plot gdp_plot
gdp_plot
```

Let's center the title.

```{r}
# save gdp_plot with centered title
gdp_plot_center <- gdp_plot +
theme( plot.title = element_text(hjust = 0.5, face = "bold" ) )

gdp_plot_center
```

Let's add the rectangles.

```{r}
gdp_plot_center +
geom_rect( data=recs.mat, aes( NULL, NULL, xmin = start, xmax = end ), ymin = 0, ymax = 200)
```

Yes, we messed up the order of the arguments in this last one.
But hey, the saving plots feature worked.
Let's change the `alpha` argument to make transparent rectangles (I should've done it from the beginning)

```{r}
p <- gdp_plot_center +
geom_rect( data=recs.mat, aes( NULL, NULL, xmin = start, xmax = end ), ymin = 0, ymax = 200, alpha = .3 )

# plot
p
```

## Themes

We can see that customizing graphs is `ggplot2`'s strong suit, there is even a package with different themes to make them even more beautiful.
Let's try it.

```{r}
library(ggthemes)

# theme bw
ggplot(data = y.df, aes( x = dates, y = GDP ) ) +
	theme_bw() +
	theme( plot.title = element_text(hjust = 0.5, face = "bold" ) ) +
	ggtitle( "Cronologia Trimestral dos Ciclos de Negócios Brasileiros" ) +
	xlab("") + ylab("") +
	geom_line( color = "blue", size=1 )
```

And now, the long waited `theme_economist()`:

```{r}
# theme economist
ggplot(data = y.df, aes( x = dates, y = GDP ) ) +
	theme_economist() +
	ggtitle( "Cronologia Trimestral dos Ciclos de Negócios Brasileiros" ) +
	xlab("") + ylab("") +
	geom_line( color = "blue", size=1 )
```

And yes, the themes work with saved objects

```{r}
gdp_plot
gdp_plot + theme_bw()
gdp_plot + theme_economist()
```

## Final Remarks

`ggplot2` is a *fenomenal* package once you jump through the initial hurdles.
The problem with `ggplot2` is that by having its own grammar the novice `R` user will not learn the `R` syntax, and the more experienced user will not benefit from knowing the `R` syntax.
Those problems arise because the package is its own animal, but also because of this it is almost infinitely customizable.
It should definitely be the choice if you intend to make beautiful polished plots.
If you just want to make a quick plot to visualize some data and forget about the plot, your are better served with the `base` plots.

But again, therein lies another problem, you will use two different *GRAMMARS* to make plots.
The `base` to make your quick and dirty plots.
And `ggplot`'s to make a finished product.
That is a question that tears me apart, and honestly I cannot decide if the tradeoff is worth.
But hey, I'm using the package, am I not?
So maybe, I have my answer.
