---
title: "Making a wrapper to Import and Plot Data"
description: |
  To make research reproducible, we have to automate the way `R` performs tasks.
categories:
  - R
  - R-base
  - GDP
  - PIB
  - plot
  - wrapper
  - reproducible-research
author:
  - name: "Paulo Ferreira Naibert"
    url: https://github.com/pfnaibert/
repository_url: https://github.com/pfnaibert/pfnaibert.github.io
date: "2020-08-18"
draft: false
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

#### Last Updated `r Sys.Date()`

In my [last post](https://pfnaibert.github.io/posts/2020-08-17-plot-base-gdp), I showed how to import a `.csv` file and plot the data.
This, entailed following the steps outlined in the post.
In this post, I will make a [wrapper function](https://en.wikipedia.org/wiki/Wrapper_function) to automate the importing and formating.

# Import Wrapper

Remembering that I already downloaded the data and saved the only column that I will need in a `.csv` file.
Our wrapper function is:

```{r}
import_gdp <- function(filename)
{

# import csv
data  <- read.csv(filename)

# delete last 2 lines
data  <- head(data, -2)

# save dates
dates <- data[,1]

# delete first row
data  <- data[,-1]

# rearrange strings in dates
dates <- gsub(" T", ":Q", dates)

# give names to values
names(data) <- dates

# `ts` object
data.ts <- ts(data, start=c(1980, 1), end=c(2014, 3), freq=4 )

return(data.ts)
}
```

And I can use `import_gdp` to import the file and format it as I like it:

```{r}
# check loaded objects
ls()

# import data
y <- import_gdp("../../data/PIB_BASE_2000.csv")

# quick visualization
head(y)

# print whole series
y

# print series time
time(y)
```

## Extensions
On this function I could have created variables to pass options about the dates and the frequency of data.
We won't do this today, but I do intend to make a following post about how to scrape the brazilian GDP data from the [SIDRA-IBGE site](https://sidra.ibge.gov.br/home/cnt/brasil).


# Plot Wrapper

Also, I can make a wrapper for the plot script.

```{r}
plot_gdp <- function(data, rec_dates)
{

# blank plot
plot(NULL,
 main = "Cronologia Trimestral dos Ciclos de Negócios Brasileiros",
 xlim = c(1980, 2015), ylim = c(60,200),
 xlab = "", ylab = "")

# recession rectangles
for(i in 1:(length(rec_dates)/2) ) rect(rec_dates[1+2*(i-1)], 1, rec_dates[2*i], 250, col="gray", border=T)

# gdp line
lines(data, col="blue", lwd=1.5)
}
```

The `plot_gdp()` function uses the `rec_dates` input to make the recession rectangles, so I must create a vector with recession limit dates.
We use [this document](https://portalibre.fgv.br/sites/default/files/2020-06/comunicado-do-comite-de-datacao-de-ciclos-economicos-29_06_2020-1.pdf) to get the dates.

```{r}
# fun to transform dates into numeric
qtr2num <- function(data) return( as.numeric(substring(data, 1, 4)) - 1/4 + as.numeric( substring(data, 7) )/4 )
```

```{r}
# recession limits in numeric dates

recs <- c( "1981:Q1", "1983:Q1",
 "1987:Q3", "1988:Q3",
 "1989:Q3", "1992:Q1",
 "1995:Q2", "1995:Q3",
 "1998:Q1", "1999:Q1",
 "2001:Q2", "2001:Q4",
 "2003:Q1", "2003:Q2",
 "2008:Q4", "2009:Q1" )

saveRDS(recs, "../../data/gdp-recessions.rds")

newrecs <- qtr2num(recs)
```

Now I can make the plot with a single line:

```{r}
# graph with wrapper
plot_gdp(y, newrecs)
```

Beacause the `rec_dates` argument generalizes the recession limit dates, I can make expansion (instead of recession) rectangles.
Again, we use [this document](https://portalibre.fgv.br/sites/default/files/2020-06/comunicado-do-comite-de-datacao-de-ciclos-economicos-29_06_2020-1.pdf) to get the dates.

```{r}
# expansion limits in numeric dates
exps <- c( "1983:Q2", "1987:Q2",
"1989:Q1", "1989:Q2",
"1992:Q2", "1995:Q1",
"1995:Q4", "1997:Q4",
"1999:Q2", "2001:Q1",
"2002:Q1", "2002:Q4",
"2003:Q3", "2008:Q3",
"2009:Q2", "2014:Q1" )

saveRDS(recs, "../../data/gdp-expansions.rds")

newexps <- qtr2num(exps)

# graph with wrapper
plot_gdp(y, newexps)
```

# Error Catching
Notice that I didn't have any error catching functions, I did this because that would be beyond the scope of the post.
If my goal was to make a more general function, I would absolutely have to include error catching lines.

# Final Remarks
To make research reproducible, we have to automate the way `R` executes tasks.
In this post I do this via wrappers.
