---
title: "Brazilian GDP 2020:Q2"
description: "Data from the 2020:Q2 Brazilian GDP"
categories:
  - R
  - ggplot2
  - dygraphs
  - scraping
  - webscraping
  - reproducible-report
  - reproducible-research
  - PIB
  - GDP
author:
  - name: "Paulo Ferreira Naibert"
    url: https://github.com/pfnaibert/
repository_url: https://github.com/pfnaibert/pfnaibert.github.io
draft: false
date: 2020-09-01
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
# knitr::opts_chunk$set(echo = TRUE, eval = TRUE, error=T, warning=T)
knitr::opts_chunk$set(fig.width=8, fig.height=4)
```

#### Last Updated `r Sys.Date()`

In this post, we will do a not so deep dive in the IBGE release of the 2020:Q2 GDP results.
As usual, my source code for the functions can be found [here](https://github.com/pfnaibert/pfnaibert.github.io/blob/master/R/funs-macro.R).

First, let's load up our functions and other libraries.

```{r, error=T, warning=T, message=T}
# source functions
source("../../R/funs-macro.R")

# load libraries
library(reshape2)
library(ggplot2)
library(dygraphs)
```

I already downloaded the data from the SIDRA website using the functions, from the `funs-macro.R` file.

```{r, eval=F}
gdp.dl()
gdp.save.rets()
gdp.save.levels()
```

Now, we load the data and make some transformations.

```{r}
# LOAD DATA

# levels
gdp.nominal <- readRDS("../../data/gdp-nominal.rds")
gdp.index   <- readRDS("../../data/gdp-index-NSA.rds")
gdp.real    <- readRDS("../../data/gdp-real-NSA.rds")

# rets
gdp.ret4    <- readRDS("../../data/gdp-ret4.rds")
gdp.ret1    <- readRDS("../../data/gdp-ret1.rds")
gdp.retsum4 <- readRDS("../../data/gdp-retsum4.rds")
gdp.retyear <- readRDS("../../data/gdp-retyear.rds")
```

```{r}
# data as ts (for dygraphs)
gdp.nominal.ts <- df2ts(gdp.nominal)
gdp.index.ts   <- df2ts(gdp.index)
gdp.real.ts    <- df2ts(gdp.real)
gdp.ret4.ts    <- df2ts(gdp.ret4)
gdp.ret1.ts    <- df2ts(gdp.ret1)
gdp.retsum4.ts <- df2ts(gdp.retsum4)
gdp.retyear.ts <- df2ts(gdp.retyear)
```

```{r}
# only GDPs
gdps <- data.frame( gdp.nominal[, c(1,2,3)],  "nominal"=gdp.nominal[, "GDP"]/1000, "index"=gdp.index[, "GDP"], "real"=gdp.real[, "GDP"]/1000 )
gdps$nominal.sum4 <- sum4(gdps$nominal)
gdps$real.sum4    <- sum4(gdps$real)
gdps$index.sum4   <- sum4(gdps$index)/4
gdps.ts <- df2ts(gdps)
```

## GGPLOT Functions

Because I will make a lot of the same plot, but using different data as input, I decided to make some `ggplot` wraper functions.

```{r}
# function to ggplot Growth by SECTORS
ggplot.last <- function(data, vars)
{
# TODO: ERROR CATCHING

newdata <- melt( sort( tail( data[, vars], 1 ), decreasing=T ) )

# plot
ggplot(newdata, aes( x = variable, y = value, fill = value > 0 ) ) +
geom_bar( stat="identity" ) +
scale_fill_manual(values=c("red", "blue") ) +
guides(fill=FALSE) +
geom_hline(yintercept=0 ) +
geom_text( aes(label=value, vjust=ifelse(value > 0, "bottom", "top") ) )+
theme(axis.text.x = element_text(color="black", size=10, angle=90) ) +
xlab("") + ylab("")
}
```

```{r}
# function to ggplot Growth by SECTORS
ggplot.growth <- function(data, ndates, vars)
{
# TODO: ERROR CATCHING

newdata      <- tail(data[, vars], ndates)
newdata$date <- date2qtr( newdata$date )
newdata      <- melt( newdata, id="date" )

#plot
ggplot(newdata, aes( x = date, y = value, group=variable, color=variable) ) +
geom_line( size=1 ) +
geom_hline(yintercept=0 ) +
theme(axis.text.x = element_text(color="black", size=11) ) +
xlab("") + ylab("")
}
```

```{r}
# vars for ggplot.last
cond.vars <- !(colnames(gdp.ret4) %in% c("date", "year", "qtr", "IND", "SER", "VAB", "TAX", "C", "FBKF", "M", "X", "G"))
vars1     <- colnames(gdp.ret4[, cond.vars])
```

## GDP GROWTH

### T/T-1 (With Seasonal Adjustment)

```{r}
ggplot.last(gdp.ret1, vars1)

ggplot.growth(gdp.ret1, ndates=8, vars=c("date", "GDP", "AGR", "IND", "SER"))

dygraph(window(gdp.ret1.ts[,c("GDP", "AGR", "IND", "SER")], start=c(2018,3) ), main = "GDP by Sectors") %>%
dyAxis("x", rangePad = 10, drawGrid = FALSE) %>%
dyLimit( as.numeric(0))


ggplot.growth(gdp.ret1, ndates=8, vars=c("date", "GDP", "C", "G", "FBKF", "X", "M"))

dygraph(window(gdp.ret1.ts[,c("GDP", "C", "G", "FBKF", "X", "M")], start=c(2018,3) ), main = "GDP by Demand Variables") %>%
dyAxis("x", rangePad = 10, drawGrid = FALSE) %>%
dyLimit( as.numeric(0))
```

#### T/T-4

```{r}
ggplot.last(gdp.ret4, vars1)

ggplot.growth(gdp.ret4, ndate=8, vars=c("date", "GDP", "AGR", "IND", "SER", "TAX"))

dygraph(window(gdp.ret4.ts[,c("GDP", "AGR", "IND", "SER", "TAX")], start=c(2018,3) ), main = "GDP by Sectors") %>%
dyAxis("x", rangePad = 10, drawGrid = FALSE) %>%
dyLimit( as.numeric(0))

ggplot.growth(gdp.ret4, ndate=8, vars=c("date", "GDP", "C", "G", "FBKF", "X", "M"))

dygraph(window(gdp.ret4.ts[,c("GDP", "C", "G", "FBKF", "X", "M")], start=c(2018,3) ), main = "GDP by Demand Variables") %>%
dyAxis("x", rangePad = 10, drawGrid = FALSE) %>%
dyLimit( as.numeric(0))
```

#### AC4Q T/T-4

```{r}
ggplot.last(gdp.retsum4, vars1)

ggplot.growth(gdp.retsum4, ndate=8, vars=c("date", "GDP", "AGR", "IND", "SER", "TAX"))

dygraph(window(gdp.retsum4.ts[,c("GDP", "AGR", "IND", "SER", "TAX")], start=c(2018,3) ), main = "GDP by Sectors") %>%
dyAxis("x", rangePad = 10, drawGrid = FALSE) %>%
dyLimit( as.numeric(0))


ggplot.growth(gdp.retsum4, ndate=8, vars=c("date", "GDP", "C", "G", "FBKF", "X", "M"))

dygraph(window(gdp.retsum4.ts[,c("GDP", "C", "G", "FBKF", "X", "M")], start=c(2018,3) ), main = "GDP by Demand Variables") %>%
dyAxis("x", rangePad = 10, drawGrid = FALSE) %>%
dyLimit( as.numeric(0))
```

## GDP GROWTH TABLES

### T/T-1 (With Seasonal Adjustment)

```{r}
paged_table(gdp.table(gdp.ret1, type="ret1", digits=1 ), list(rows.print=12) )
```

### T/T-4

```{r}
paged_table(gdp.table(gdp.ret4, type="ret4", digits=1), list(rows.print=12) )
```

### AC4Q T/T-4

```{r}
paged_table(gdp.table(gdp.retsum4, type="ret4", digits=1), list(rows.print=12) )
```

## GDP LEVELS

### NOMINAL

```{r}
dygraph(gdps.ts[,"nominal"], main = "Nominal GDP (Billions of Current BRL)") %>%
dySeries(label = "Nominal GDP") %>%
dyAxis("x", rangePad = 10, drawGrid = FALSE) %>%
dyLimit( as.numeric(gdps.ts[NROW(gdps), "nominal"]) )


dygraph(gdps.ts[,"nominal.sum4"], main = "Nominal GDP AC4Q (Billions of Current BRL)") %>%
dySeries(label = "Nominal GDP") %>%
dyAxis("x", rangePad = 10, drawGrid = FALSE) %>%
dyLimit( as.numeric(gdps.ts[NROW(gdps), "nominal.sum4"]) )
```

### REAL

```{r}
dygraph(gdps.ts[,"index"], main = "GDP INDEX (1995=100)") %>%
dySeries(label = "GDP") %>%
dyAxis("x", rangePad = 10, drawGrid = FALSE) %>%
dyLimit( as.numeric(gdps.ts[NROW(gdps), "index"]) )

dygraph(gdps.ts[,"index.sum4"], main = "GDP INDEX Mean of 4 Quarters (1995=100)") %>%
dySeries(label = "GDP") %>%
dyAxis("x", rangePad = 10, drawGrid = FALSE) %>%
dyLimit( as.numeric(gdps.ts[NROW(gdps), "index.sum4"]) )
```

## GDP LEVELS TABLES

### NOMINAL (Millions of Current BRL)

```{r}
paged_table( gdp.table(gdp.nominal, type="level" , period="qtr", digits=0), list(rows.print=12) )
```
[SIDRA TABLE](https://sidra.ibge.gov.br/tabela/1846#resultado)

### REAL (Milliions of 1995 Chained BRL)

```{r}
paged_table( gdp.table(gdp.real, type="ret4" , period="qtr", digits=0), list(rows.print=12) )
```
[SIDRA TABLE](https://sidra.ibge.gov.br/tabela/6612#resultado)

### INDEX (Chained Index)

```{r}
paged_table( gdp.table(gdp.index, type="ret4" , period="qtr", digits=2), list(rows.print=12) )
```
[SIDRA TABLE](https://sidra.ibge.gov.br/tabela/1620#resultado)

## NEXT RELEASES FROM IBGE

|Reference Date | Publishing Date
|:--- |:--- |
| 2020:Q3 |	2020-12-03 |


## Useful Links
- [SCNT](https://www.ibge.gov.br/estatisticas/economicas/contas-nacionais/9300-contas-nacionais-trimestrais.html?=&t=o-que-e)
- [SCNT-releases](https://biblioteca.ibge.gov.br/index.php/biblioteca-catalogo?view=detalhes&id=72121)
- [2020:Q1 IBGE release](https://biblioteca.ibge.gov.br/visualizacao/periodicos/2121/cnt_2020_1tri.pdf)
- [2020:Q2 IBGE release](https://biblioteca.ibge.gov.br/visualizacao/periodicos/2121/cnt_2020_2tri.pdf)
- [SIDRA-CNT](https://sidra.ibge.gov.br/home/cnt/brasil)
- [sthda-ggplot](http://www.sthda.com/english/wiki/ggplot2-essentials)
- [R-cookbook plots](http://www.cookbook-r.com/Graphs/)

# THANK YOU FOR READING!
