---
title: "Scraping Brazilian GDP data from the SIDRA-IBGE website"
description: "Scraping Brazilian GDP data from the SIDRA-IBGE website"
categories:
  - R
  - ggplot2
  - dygraphs
  - scraping
  - webscraping
  - reproducible-report
  - reproducible-research
  - PIB
  - GDP
author:
  - name: "Paulo Ferreira Naibert"
    url: https://github.com/pfnaibert/
repository_url: https://github.com/pfnaibert/pfnaibert.github.io
draft: false
date: 2020-08-24
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, error = FALSE, warning = FALSE)
knitr::opts_chunk$set(fig.width=8, fig.height=4)
```

#### Last Updated `r Sys.Date()`

In this post, we will scrape the brazilian gdp data from the [SIDRA-IBGE](https://sidra.ibge.gov.br/) website.
Here, I use my own functions (based on other codes) to download the sidra data, but be aware that theres is a CRAN package for that [here](https://cran.r-project.org/web/packages/sidrar/index.html).
My source code for the functions can be found [here](https://github.com/pfnaibert/pfnaibert.github.io/blob/master/R/funs-macro.R).
Also, there is another post [here](https://www.jfcostasantos.com/post/2015-07-23-r-rmarkdown/) that makes a lot of the things we will be doing.

Ok, disclaimers are out of the way.
First, let's load up our functions and other libraries.

```{r}
# source functions
source("./myfuns.R")

# load libraries
library(dplyr)
library(reshape2)
library(lubridate)
library(dygraphs)
library(ggplot2)
library(ggthemes)
```

The `gdp.dl.csv()` function, downloads the data from the website and saves a `.csv`.
The `gdp.save.levels()` and `gdp.save.rets()` functions load the `.csv` file, using the paths used in `gdp.dl.csv()`, transforms the data using the `gdp.csv2df()` function and saves a `data.frame` file with the `.rds` format.
I separated the functions that download and trassform the data to avoid connecting to the internet every time I wanted to edit the `.csv` file.
Also, the function `df2ts()` transforms the `data.frame` into a `ts` object because `paged_table()` and `ggplot()` do not accept `ts` objects, whereas `dygraphs` only accepts time series objects.

```{r, eval=F}
# download gdp
gdp.dl.csv()

# transform and save data
gdp.save.levels()
gdp.save.rets()
```

Now, let's load our data.

```{r}
# LOAD DATA

# levels
gdp.nominal  <- readRDS("../../data/gdp-nominal.rds")
gdp.real.NSA <- readRDS("../../data/gdp-real-NSA.rds")
gdp.real.SA  <- readRDS("../../data/gdp-real-SA.rds")

# rets
gdp.ret4    <- readRDS("../../data/gdp-ret4.rds")
gdp.ret1    <- readRDS("../../data/gdp-ret1.rds")
gdp.retsum4 <- readRDS("../../data/gdp-retsum4.rds")
gdp.retyear <- readRDS("../../data/gdp-retyear.rds")

# data as ts (for dygraphs)
gdp.nominal.ts  <- df2ts(gdp.nominal)
gdp.real.NSA.ts <- df2ts(gdp.real.NSA)
gdp.real.SA.ts  <- df2ts(gdp.real.SA)
gdp.ret4.ts     <- df2ts(gdp.ret4)
gdp.ret1.ts     <- df2ts(gdp.ret1)
gdp.retsum4.ts  <- df2ts(gdp.retsum4)
gdp.retyear.ts  <- df2ts(gdp.retyear)

# only GDPs
gdp.df <- data.frame( gdp.nominal[, c(1,2,3)],  "nominal"=gdp.nominal[, "GDP"], "real.NSA"=gdp.real.NSA[, "GDP"], "real.SA"=gdp.real.SA[, "GDP"],
					 "ret1"=gdp.ret1[, "GDP"], "ret4"=gdp.ret4[, "GDP"], "retsum4"=gdp.retsum4[, "GDP"] )
gdp.df$sum4.nomi    <- sum4(gdp.df$nominal)
gdp.df$sum4.real    <- sum4(gdp.df$real.NSA)
gdp.ts <- df2ts(gdp.df)
```

```{r}
# show gdp.df
paged_table(gdp.df)
```

Cool, the data is loaded, let's make a table with the annual gdp data starting in 2014.

```{r}
tmp <- filter(gdp.df, date > "2014-01-01" & month(date) == "12" )
mat <- cbind( "Billions of current BRL"=tmp$sum4.nomi/1000, "Billions of 1995 BRL"=tmp$sum4.real/1000, "AC 4Q T/T-4"=tmp$retsum4)
rownames(mat) <- substring( date2qtr(tmp$date), 1, 4)
kable( t( apply( mat, 2, rev) ), digits=2, caption = "GDP by Year")
```

Now, let's make a table with quarterly gdp data starting from 2019.

```{r}
tmp <- filter(gdp.df, date > "2019-01-01" )
mat <- cbind( "Billions of current BRL"=tmp$nominal/1000, "T/T-1"=tmp$ret1, "T/T-4"=tmp$ret4, "AC 4Q T/T-4"=tmp$retsum4)
rownames(mat) <- date2qtr(tmp$date)
kable( t( apply( mat, 2, rev) ), digits=2, caption = "GDP by Quarter")
```

Tables are ready, let's graph the GDP percent change for better visualization in `ggplot`:

```{r}
gdp.df %>%
filter(date > "2014-01-01") %>%
select(date, ret1, ret4, retsum4) %>%
melt("date") %>%
ggplot( aes( x = date, y = value, fill = variable) ) +
geom_bar( stat="identity", position="dodge" ) +
scale_fill_discrete(name=" ",
					breaks=c("ret1", "ret4", "retsum4"),
					labels=c("T/T-1", "T/T-4", "T/T-4 AC4Q") ) +
xlab("") + ylab("") +
ggtitle( "Brazilian GDP %Change" ) +
theme( plot.title = element_text(hjust = 0.5, face = "bold" ) ) +
theme_economist()
```

and interactive version with `dygraphs`:

```{r}
tmp <- window(gdp.ts, start=c(2014,1) )
dygraph( tmp[, c("ret1", "ret4", "retsum4") ], main="Brazilian GDP %Change" ) %>%
dySeries("ret1", label = "T/T-1") %>%
dySeries("ret4", label = "T/T-4") %>%
dySeries("retsum4", label = "AC 4Q T/T-4") %>%
dyAxis("x", rangePad = 20, drawGrid = FALSE) %>%
dyBarChart()
```

Now, let's graph the level of the GDP using the Index of chained 1995 BRL, seasonally adjusted and not seasonally adjusted.
In `ggplot2`:

```{r}
den.SA  <- mean(gdp.df$real.SA[year(gdp.df$date)=="2014"] )
den.NSA <- mean(gdp.df$real.NSA[year(gdp.df$date)=="2014"] )

gdp.df %>%
select(date, real.SA, real.NSA) %>%
mutate(real.SA = 100*real.SA/den.SA, real.NSA = 100*real.NSA/den.NSA ) %>%
melt(id = "date") %>%
ggplot( aes(x = date, y = value) ) +
  ggtitle( "GDP Index of 1995 Chained BRL (2014=100)" ) +
  geom_line(aes(color = variable), size = 1) +
  scale_color_manual(name = "", values = c("blue", "red"), labels=c("Seasonal Adj.", "Not Seasonal Adj." ) ) +
  theme( plot.title = element_text(hjust = 0.5, face = "bold" ) ) +
  xlab("") + ylab("") +
  theme_economist()
```

And with `dygraphs`:

```{r}
tmp <- cbind( "real.SA"=normalize.yr( gdp.ts[, "real.SA"], 2014), "real.NSA"=normalize.yr(gdp.ts[, "real.NSA"], 2014 ) )
dygraph(tmp, main="GDP Index of 1995 Chained BRL (2014=100)" ) %>%
dyAxis("x", rangePad = 10, drawGrid = FALSE) %>%
dyEvent("2002-01-1", "Lula",  labelLoc = "bottom") %>%
dyEvent("2010-01-1", "Dilma", labelLoc = "bottom") %>%
dyEvent("2016-09-1", "Temer", labelLoc = "bottom") %>%
dyEvent("2019-01-1", "Bolsonaro", labelLoc = "bottom") %>%
dyEvent("2008-10-1", "Subprime", labelLoc = "bottom", color="red") %>%
dyEvent("2020-03-01", "Corona", labelLoc = "bottom", color="red") %>%
dyLimit( max(tmp), "Peak", labelLoc = "left", color="blue" ) %>%
dySeries("real.NSA", label = "NSA") %>%
dySeries("real.SA", label = "SA")
```

We could (does not mean that we should) also normalize by an specific quarter.

```{r}
den.SA  <- gdp.df$real.SA[gdp.df$date=="2013-09-30"]
den.NSA <- gdp.df$real.NSA[gdp.df$date=="2013-09-30"]

gdp.df %>%
select(date, real.SA, real.NSA) %>%
mutate(real.SA = 100*real.SA/den.SA, real.NSA = 100*real.NSA/den.NSA ) %>%
melt(id = "date") %>%
ggplot( aes(x = date, y = value) ) +
  ggtitle( "GDP Index of 1995 Chained BRL (2013:Q3=100)" ) +
  geom_line(aes(color = variable), size = 1) +
  scale_color_manual(name = "", values = c("blue", "red"), labels=c("Seasonal Adj.", "Not Seasonal Adj." ) ) +
  theme( plot.title = element_text(hjust = 0.5, face = "bold" ) ) +
  xlab("") + ylab("") +
  theme_economist()
```

```{r}
tmp <- cbind( "real.SA"=normalize( gdp.ts[, "real.SA"], c(2013, 3)), "real.NSA"=normalize(gdp.ts[, "real.NSA"], c(2013,3) ) )
dygraph(tmp, main="GDP Index of 1995 Chained BRL, (2013:Q3=100)" ) %>%
dyAxis("x", rangePad = 10, drawGrid = FALSE) %>%
dyEvent("2002-01-1", "Lula",  labelLoc = "bottom") %>%
dyEvent("2010-01-1", "Dilma", labelLoc = "bottom") %>%
dyEvent("2016-09-1", "Temer", labelLoc = "bottom") %>%
dyEvent("2019-01-1", "Bolsonaro", labelLoc = "bottom") %>%
dyEvent("2008-10-1", "Subprime", labelLoc = "bottom", color="red") %>%
dyEvent("2020-03-01", "Corona", labelLoc = "bottom", color="red") %>%
dyLimit( max(tmp), "Peak", labelLoc = "left", color="blue" ) %>%
dySeries("real.NSA", label = "NSA") %>%
dySeries("real.SA", label = "SA")
```

So that is our little report of the brazilian GDP series.

## Extensions
On future posts I intend to break it up by sectors and components.

## Automation and cron jobs
Also, we can run the `gdp.dl.csv()` and  the `gdp.save.levels()` and `gdp.save.rets()` functions separately on an executable script using cron jobs ([more details here](https://en.wikipedia.org/wiki/Cron)) to autmate the scraping part.
The rest of the script is completely reusable to automate reporting.

## Future dates
Finally, [here](https://sidra.ibge.gov.br/home/cnt/brasil) we can find out when the next research will be published:

| Reference Date | Publishing Date |
| ---- | ---- |
|  2020:Q2       | "2020-09-01"    |
|  2020:Q3       | "2020-12-03"    |

Stay tuned.

# Thank you for reading!
